---
title: "Data 608 - HW 1"
author: "Georgia Galanopoulos"
date: "February 11, 2018"
output: 
  html_document:
    keep_md: true
---
```{r}
library(knitr)
library(dplyr)
library(FSA)
library(ggplot2)
```

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)
```

And let's preview this data:

```{r}
head(inc)
summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}
# Structure of inc
str(inc)

# Number of NAs for each column
sapply(inc, function(x){sum(length(which(is.na(x))))})

# Subset for NA values
NAinc = inc[is.na(inc$Employees),]
summary(NAinc)
kable(NAinc)

# Correlation Matrix
noNA = inc[complete.cases(inc), ]
round(cor(noNA[,sapply(noNA, is.numeric)]),2)

# Viewing Growth Rate by State
grsta = data.frame(Summarize(Growth_Rate ~ State, data=noNA))
grsta[order(grsta$mean, decreasing = T),]
```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r, fig.height = 11, fig.width = 5}
ggplot(grsta, aes(x = reorder(State, n), y = n)) + 
  geom_histogram(stat="identity") + 
  coord_flip() + 
  ylab("Number of Companies") +
  xlab("State")
```

California, Texas and New York were the top three states with the highest number of companies per state, with California in the lead by a large margin.

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r, fig.height = 5, fig.width = 11}
# Picking the third company
state = grsta[order(grsta$n, decreasing = T),][3,"State"]
onlystate = inc[which(inc$State== state), ]

# Checking cases
which(complete.cases(onlystate)!= T)

## Graph of Employment by Industry in NY
# Remove extreme outliers
outliers = onlystate$Employees %in% boxplot.stats(onlystate$Employees, coef = 1.5)$out
outclean = onlystate[!outliers, ]

# Boxplots to show range
ggplot(outclean, aes(x = reorder(Industry, Employees, FUN = median, na.rm = TRUE), y = Employees)) + 
  geom_boxplot()+ 
  ylab("Number of Employees") +
  xlab("Industry")+
  coord_flip()
```

Extreme outliers have been removed, but the those within 2 standard deviations remained. Since the majority of the industries (especially in the lower end) are skewed, the median was chosen as a more appropriate parameter estimate of the number of employees per industry. Energy and Financial Services were the top two fields with the highest employees. Government Services and Retail were the bottom two. Industries with impressive ranges of employees were Software and Financial Services.


## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r, fig.height = 5, fig.width = 11}
# Add Revenue-Per-Employee variable
RPE = noNA %>% mutate(RevPerEmp = Revenue/Employees)
# Remove extreme outliers
RPEoutliers = RPE$RevPerEmp %in% boxplot.stats(RPE$RevPerEmp, coef = 1.5)$out
RPEoutclean = RPE[!RPEoutliers, ]

# Revenue/Employee vs Industry
ggplot(RPEoutclean, aes(x = reorder(Industry, RevPerEmp, FUN = mean), y = RevPerEmp)) + 
  geom_boxplot(outlier.shape = NA) +
  ylab("Revenue per Employee") + 
  xlab("Industry")+
  coord_flip()+
  stat_summary(fun.y=mean, colour="blue", geom="point", show_guide = FALSE)
```

The boxplots here were orderd by mean rather than median because they majority of them appear to be normally distributed. It appears that the "Computer Hardware" industry generates the most revenue per employee, followed by "Logistics & Transporation" and "Retail". The majority of the industries have an impressive range, though. Human Resources, despite having one of the lowest means, has has the largest range of the bottom five industries.
