---
title: "Data 608 - HW 1"
author: "Georgia Galanopoulos"
date: "February 4, 2018"
output: 
  html_document:
    keep_md: true
---
```{r}
library(knitr)
library(dplyr)
library(FSA)
library(ggplot2)
```

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)
```

And lets preview this data:

```{r}
head(inc)
summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}
# Structure of inc
str(inc)

# Number of NAs for each column
sapply(inc, function(x){sum(length(which(is.na(x))))})

# Subset for NA values
kable(inc[is.na(inc$Employees),])

# Correlation Matrix
noNA = inc[complete.cases(inc), ]
round(cor(noNA[,sapply(noNA, is.numeric)]),2)

# Viewing Growth Rate by State
grsta = data.frame(Summarize(Growth_Rate ~ State, data=noNA))
grsta[order(grsta$mean, decreasing = T),]
```

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r, fig.height = 11, fig.width = 5}
# Number of Companies by State
ggplot(grsta, aes(x = reorder(State, n), y = n)) + 
  geom_histogram(stat="identity") + 
  coord_flip() + 
  ylab("Companies")
```

## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r, fig.height = 5, fig.width = 11}
# Picking the third company
state = grsta[order(grsta$n, decreasing = T),][3,"State"]
onlystate = inc[which(inc$State== state), ]

# Checking cases
which(complete.cases(onlystate)!= T)

# Graph of Employment by Industry in NY
ggplot(onlystate, aes(x = reorder(Industry, Employees, FUN = median), y = Employees)) + 
  # boxplots show mean and range
  geom_boxplot(outlier.shape = NA) +                                        # remove outliers
  scale_y_continuous(limits = quantile(onlystate$Employees, c(0.1, 0.9))) + # shorten y-axis
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+ 
  ylab("Number of Employees") +
  xlab("Industry")
```

The "Environmental Services" employ the most people on average, followed closely by "Energy". On the other hand, "Health", "IT Services" and "Telecommunications" have the most impressive ranges.

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r, fig.height = 5, fig.width = 11}
# Add Revenue-Per-Employee variable
RPE = noNA %>% mutate(RevPerEmp = Revenue/Employees)

# Revenue/Employee vs Industry
ggplot(RPE, aes(x = reorder(Industry, RevPerEmp, FUN = mean), y = RevPerEmp)) + 
         geom_boxplot(outlier.shape = NA) +
         ylab("Revenue per Employee") + 
         xlab("Industry") +
         scale_y_continuous(limits = quantile(RPE$RevPerEmp, c(0.1, 0.9)))  +
         theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

It appears that the "Logistics & Transporation" industry generates the most revenue per employee, followed closely by "Retail" and "Computer Hardware".
